Correcting K-correction: dark energy is based on a math error from 1930

### 2025-01-06 paper/dark_energy_corrections.tex

\begin{abstract}
We rederive the K-correction equation and discover that the version of the
equation used to compute magnitudes contains three errors. The oldest of these
errors traces back nine decades; in 1930 the equation did not correct for
spectral bandwidth stretching, followed by a rederivation in 1968 that
accounted for spectral bandwidth stretching but did not correct for time
dilation. Two new errors were introduced in 1996, one of which causes
systematic distance biases by inverting the zero-point correction term. We
conclude that properly correcting apparent magnitudes for redshift resolves two
of the preeminent mysteries in cosmology: first, dark energy is not supported
by observations of Type Ia supernovae and second, the Hubble tension is due to
the errors in the K-correction equation.
\end{abstract}

@@

The first formal derivation of K-correction was performed by
\citet{tolman1930}, but it did not account for spectral bandwidth stretching,
one of the dimming effects of redshift. An alternative K-correction was
correctly derived by \citet{desitter1934}. However, in \citet{hubble1935}, the
discrepancy was noted yet dismissed. \citet{oke1968} rederived the K-correction
equation.  They accounted for spectral bandwidth stretching, but they did not
account for time dilation, one of the other dimming effects of redshift. The
paper by \citet{kim1996}, which is the basis for modern implementations of
K-correction, extended the work of \citet{oke1968} to handle additional
cross-filter comparisons and added a term that was intended to address
zero-point corrections. This history is explored in more depth in Section
\ref{sec:history}.

The effect of using an incorrect magnitude for SnIa is that we think objects
are farther away than they really are, and the effect compounds for greater
distances. Up until \citet{riess1998}, we did not have distant enough
observations for this error to matter much for cosmological models, but by the
late 1990s, it became clear that our measurements for distance and redshift were
not linear. This realization led to a model that utilized a cosmological
constant $\Lambda$ and dark energy in order to explain the non-linear
distance-redshift graph. We will show in Section \ref{sec:tension} that
correcting the flaw with K-correction leads to a linear graph that does not
need to rely on a cosmological constant.

\citet{planck2015} used an alternative technique to compute the Hubble constant
$H_0$ that is based on measurements of the cosmic microwave background (CMB).
CMB-based computations of $H_0$ has made it evident that something was missing
with our understanding of cosmology because this technique produces a
significantly different value than when the constant is derived using SnIa.  As
we will discuss in Section \ref{sec:tension}, fixing the error with
K-correction produces a measurement of the Hubble constant that is compatible
with models based on the CMB.

@@

\begin{figure}
  \plotone{lum_distance_vs_redshift.pdf}
  \caption{The relationship between luminosity distance and redshift for two
  treatments of magnitude data. The displayed points are roughly a third of the
  values in the full SnIa dataset published in \citet{abbott2024}, selected
  evenly to aid visibility. The $k(z) = 1$ treatment, which represents the
  uncorrected values, is clearly non-linear while the $k(z) = 1 + z$ treatment
  which fixes the K-correction redshift error, appears to be linear.}
  \label{fig:mu_distance_vs_redshift}
\end{figure}

@@

However, redshifting light dims an observation in three ways:

@@

  \item The bandwidth stretches. If a rest frame observes photons
  between the wavelengths of 400nm and 401nm, then a redshift of $z = 1$
  stretches the wavelength from 800nm to 802nm.

  \item Cosmological time dilation reduces the rate at which photons arrive.
\end{itemize}

To complicate matters, we use the $B$ band magnitude to calculate luminosity
distance. We want to know how bright a supernova appears for wavelengths around
445nm as if no redshift had occurred. However, we typically need to observe the
supernova with a filter that is sensitive to longer wavelengths, such as the
$i$ band filter which is sensitive to wavelengths from 700nm-850nm
\citep{flaugher2015}. The K-correction formula allows us to take a magnitude
measured in an observation filter $y$ and compute the magnitude in the target
filter $y$.

The first mathematical treatments of K-correction was performed by
\citet{tolman1930}. However, when Tolman made his derivation, he did not
consider the effects of a spectrum that is stretched due to redshift. See
the top two panels of Figure \ref{fig:k-example} for an example of this effect.

@@

A year later, \citet{hubble1935} published a similar set of calculations for
K-correction, but these equations used $(1 + z)^2$ instead of the $(1 + z)^3$
correction term used by de Sitter. They started their derivation by copying the
incorrect equation from 1930. After their derivation, they noted:

@@

The Hubble K-correction with the incorrect correction term have been used ever
since.

By \citet{oke1968}, the two factors of $(1 + z)$ were attributed to the change
in energy and to the spectral bandwidth elongation, which leaves time dilation
as the factor that was omitted. A graph that demonstrates why it is essential to
correct for both the spectra bandwidth stretching and time dilation is presented
in Figure \ref{fig:k-example}.

The modern treatment of K-correction is based on the work of \citet{kim1996}.
This work extended the calculations of K-correction to extend to filters beyond
the blue $B$ and visible $V$ bands. It also introduced a term that deals with
the zero-point for the actual filters. In the modern day, filters measure the
photon flux as opposed to the energy flux. Historically, bolometric devices
would measure the energy flux, but modern charge-coupled device (CCD) cameras
effectively measure the photon flux.

@@

With modern CCD cameras, a telescope observation consists of a single value
$\mathcal{F}_x$ erg/s, which represents the energy collected in filter $x$ per
second. A summary of how this works is provided by \citet{lesser2015}.
Photogenerated electrons are collected in potential wells, which means that the
measured energy is proportional to the number of photons. However, we need to
calculate the expected number of photons collected by using a spectral energy
density function $F$. The value $F(\lambda)$ gives the amount of energy
collected by a bolometric device for the wavelength $\lambda$.

@@

However, this equation is misleading and error prone. We will want to use it to
help calculate the amount of flux in an observation filter at the redshifted
wavelength $\lambda \times (1 + z)$. In other words, we want to produce the
photon density function $R'$ that is the redshifted version of $F'$.
Redshifting the photon density does two things:

\begin{itemize}
  \item The stretching of space increases the distance between photons while
  they are traveling. This phenomenon appears to an observer like time
  dilation, although cosmological time dilation is due to a different mechanism
  than relativistic time dilation. This effect reduces the photon arrival rate
  by a factor of $1/(1+z)$.

  In order to account for cosmological time dilation, we will need to multiply
  $F'$ by $1/(1+z)$.

  \item All wavelengths are changed by a factor of $1 + z$. When we integrate
  $R'$ from wavelength $\lambda_a$ to wavelength $\lambda_b$, the values
  correspond to the wavelengths $\lambda_a / (1+z)$ to $\lambda_b / (1+z)$ in
  $F'$. We will integrate over a width of $\lambda_b - \lambda_a$, but
  $F'(\lambda / (1+z))$ will refer to values in a width of
  $(\lambda_b - \lambda_a) / (1+z)$.

  In order to account for this spectral bandwidth stretching effect, we will need
  to multiply $F'$ by a second factor of $1/(1+z)$.
\end{itemize}

@@

We deliberately do not combine all of the factors of $1+z$ together because
this form is more natural to implement in code.

@@

The limits of integration are technically from 0 to $\infty$, but these are
usually not written because the sensitivity $S(\lambda)$ is 0 for wavelengths
outside of a filter's bandpass.

@@

$P_x$ represents the zero-point for the filter $x$ on some particular
telescope. In order to use consistent magnitude values across telescopes that
have different light gathering abilities, we take the measured magnitude and
multiply it by the ratio of the standard flux rate to the flux rate for this
particular telescope and filter. For convenience, we use $P_x = -2.5
\text{log}(P'_x)$ so that we can work with flux instead of with magnitude.

Now that we have the identities in Equations \ref{eq:flux_definition} and
\ref{eq:flux2mag} we will change directions and look at the definition of
K-correction $K_{xy}$. This value allows us to make an observation in filter
$y$ and report what the magnitude would have been in filter $x$ if no redshift
occurred:

@@

Note that this substitution omits the minus ($-$) sign that we used on a
similar substitution for $P_x$.

@@

Finally, we can use Equation \ref{eq:k_substitution} to convert the flux $K'_{xy}$ back
into the magnitude $K_{xy}$:

@@

  This effect by itself can alter estimates of $H_0$ by several km s$^{-1}$
  Mpc$^{-1}$.  While we present an estimate of the Hubble constant in Figure
  \ref{fig:H0bootstrap}, we are not terribly confident in the numbers because
  we were unable to correct for the zero-point errors, or verify that these
  errors may not exist.

@@

\section{Dark energy and the Hubble tension}
\label{sec:tension}

As shown in Figure \ref{fig:mu_distance_vs_redshift}, when reported magnitudes
are corrected by adding the magnitude $-2.5\text{log}(1+z)$, there is a linear
relationship between redshift and luminosity distance. This is in accordance
with the Hubble-Lema\^{i}tre law. As opposed to the uncorrected values, there
is no visual acceleration.

@@

Since the corrected values are approximately linear, we can use them to
estimate the Hubble constant $H_0$. This is demonstrated in Figure
\ref{fig:expansion}. Furthermore, we performed a bootstrap calculation to
estimate the distribution of $H_0$ estimates in Figure \ref{fig:H0bootstrap}.

\begin{figure}
  \plotone{bootstrapped_H0.pdf}
  \caption{A histogram of $10{,}000$ bootstrap trials measuring the Hubble constant
  $H_0$. Each trial samples an absolute magnitude $M \sim Norm(-19.2334,
  0.0404)$ for SnIa based on data published by \citet{camarena2020}. It then samples,
  with replacement, a population of supernovae from the dataset published by
  \citet{abbott2024}. Finally, it uses the non-parametric linear regression
  technique described by \citet{siegel1982}. The result of the bootstrap is $H_0
  \sim Norm(65.94, 1.30^2)$.
  }
\label{fig:H0bootstrap}
\end{figure}

The value estimated here, $H_0 \sim 65.94 \pm 1.30$ is highly suspicious
because of the flipped zero-point corrections, but it is consistent with
estimates of $H_0$ that are based on the CMB. \citet{planck2020} published the
CMB value of $H_0 \sim 67.27 \pm 0.6$, and the one sigma error bars overlap. It
is interesting to note that restricting the Dark Energy Survey observations of
SnIa \citep{vincenzi2024} to only those with $z < 0.1$, which presumably is a
small enough redshift that all observations will have been made with the same
filter, gives the estimate $H_0 \sim 67.8 \pm 1.86$.

\section{Conclusions}

>>>>

\begin{abstract}
We rederive the K-correction equation and discover that the version of the
equation that is commonly used to calculate magnitudes contains three errors.
The oldest of these errors traces back nine decades; in 1930 the equation did
not correct for spectral bandwidth stretching, followed by a rederivation in
1968 that accounted for spectral bandwidth stretching but did not correct for
time dilation. Two new errors were introduced in 1996, one of which causes
systematic distance biases by inverting the zero-point correction term. We
conclude that properly correcting apparent magnitudes for redshift resolves two
of the preeminent mysteries in cosmology: first, dark energy is not supported
by observations of Type Ia supernovae and second, the Hubble tension is due to
the errors in the K-correction equation.
\end{abstract}

@@

\citet{tolman1930} published the first formal derivation of K-correction, but
he did not account for spectral bandwidth stretching, one of the dimming
effects of redshift. An alternative K-correction was correctly derived by
\citet{desitter1934}. However, in \citet{hubble1935}, the discrepancy was noted
yet dismissed. \citet{oke1968} rederived the K-correction equation where they
accounted for spectral bandwidth stretching, but they did not account for time
dilation, one of the other dimming effects of redshift. The paper by
\citet{kim1996}, which is the basis for modern implementations of K-correction,
extended the work of \citet{oke1968} to handle additional cross-filter
comparisons and added a term that was intended to address zero-point
corrections. This history is explored in more depth in Section
\ref{sec:history}.

The effect of using an incorrect magnitude for SnIa is that we think objects
are farther away than they really are, and the effect compounds for greater
distances. Up until \citet{riess1998}, we did not have distant enough
observations for this error to matter much for cosmological models, but by the
late 1990s, it became clear that our measurements for distance and redshift were
not linear. This realization led to a model that utilized a cosmological
constant $\Lambda$ and dark energy in order to explain the non-linear
distance-redshift graph. We will show in Section \ref{sec:conclusions} that
correcting the flaws with K-correction leads to a linear graph that does not
need to rely on a cosmological constant.

\citet{planck2015} used an alternative technique to compute the Hubble constant
$H_0$ that is based on measurements of the cosmic microwave background (CMB).
CMB-based computations of $H_0$ have made it evident that something is missing
with our understanding of cosmology because this technique produces a
significantly different value than when the constant is derived using SnIa.  As
we will discuss in Section \ref{sec:conclusions}, fixing the errors with
K-correction produces a measurement of the Hubble constant that is compatible
with models based on the CMB.

@@

\begin{figure}
  \plotone{lum_distance_vs_redshift.pdf}
  \caption{The relationship between luminosity distance and redshift for two
  treatments of magnitude data. The displayed points are roughly a third of the
  values in the full SnIa dataset published in \citet{abbott2024}, selected
  evenly to aid visibility. The red $k(z) = 1$ treatment, which represents the
  uncorrected values, is clearly non-linear while the blue $k(z) = 1 + z$
  treatment, which fixes the K-correction redshift error, appears to be linear.}
  \label{fig:mu_distance_vs_redshift}
\end{figure}

@@

However, redshifting dims the luminosity of an observation in three ways:

@@

  \item The bandwidth stretches. If the rest frame observes photons in the
  bandwidth from 400nm and 401nm, then a redshift of $z = 1$ stretches the
  wavelengths to the bandwidth from 800nm to 802nm. The number of photons per
  bandwidth is reduced by a factor of $1 / (1+z)$.

  \item Cosmological time dilation reduces the rate at which photons arrive.
  The number of photons that arive per second is reduced by a factor of $1 / (1+z)$.
\end{itemize}

To complicate matters, we use the $B$ band magnitude to calculate luminosity
distance. We want to know how bright a supernova appears for wavelengths around
445nm as if no redshift had occurred. However, we typically need to observe the
supernova with a filter that is sensitive to longer wavelengths, such as the
$i$ band filter which is sensitive to wavelengths from 700nm-850nm
\citep{flaugher2015}. The K-correction formula allows us to take a magnitude
measured in an observation filter $y$ and compute the magnitude in the target
filter $x$.

The first mathematical treatment of K-correction was performed by
\citet{tolman1930}. However, when Tolman made his derivation, he did not
consider the effects of a spectrum that is stretched due to redshift.

@@

A year later, \citet{hubble1935} published a similar set of calculations where they noted:

@@

\noindent However, even though they considered all three dimming effects of
redshift, they started their derivation by copying the incorrect equation from
1930. The incorrect correction term has been used ever since.

In a paper published by \citet{oke1968}, the two factors of $(1 + z)$ were
attributed to the change in energy and to the spectral bandwidth elongation,
which leaves time dilation as the factor that was omitted.

The modern treatment of K-correction is based on the work of \citet{kim1996}.
This work extended the calculations of K-correction to extend to filters beyond
the blue $B$ and visible $V$ bands. It also introduced a term that deals with
the zero-point for the actual filters. Historically, bolometric devices would
measure the energy flux, but modern charge-coupled device (CCD) cameras
effectively measure the photon flux.

@@

With modern CCD cameras, a telescope observation consists of a single value
$\mathcal{F}_y$ erg/s, which represents the energy collected per second in
filter $y$. A summary of how this works is provided by \citet{lesser2015}.
Photogenerated electrons are collected in potential wells, which means that the
measured energy is proportional to the number of photons. However, we need to
calculate the expected number of photons collected by using a spectral energy
density function $F$. The value $F(\lambda)$ gives the amount of energy
collected by a bolometric device for the wavelength $\lambda$.

@@

\noindent However, this equation is misleading and error prone. We will want to use it to
help calculate the amount of flux in an observation filter at the redshifted
wavelength $\lambda \times (1 + z)$. In other words, we want to produce the
photon density function $R'$ that is the redshifted version of $F'$.
Redshifting the photon density does two things:

\begin{itemize}
  \item All wavelengths are increased by a factor of $1 + z$. When we integrate
  $R'$ from wavelength $\lambda_a$ to wavelength $\lambda_b$, the values
  correspond to the wavelengths $\lambda_a / (1+z)$ to $\lambda_b / (1+z)$ in
  $F'$. We will integrate over a width of $\lambda_b - \lambda_a$, but
  $F'(\lambda / (1+z))$ will refer to values in a width of
  $(\lambda_b - \lambda_a) / (1+z)$.

  In order to account for this spectral bandwidth stretching effect, we will need
  to multiply $F'$ by a second factor of $1/(1+z)$.
  \item The stretching of space increases the distance between photons while
  they are traveling. This phenomenon appears to an observer like time
  dilation, although cosmological time dilation is due to a different mechanism
  than relativistic time dilation. This effect reduces the photon arrival rate
  by a factor of $1/(1+z)$.

  In order to account for cosmological time dilation, we will need to multiply
  $F'$ by $1/(1+z)$.
\end{itemize}

@@

\noindent We deliberately do not combine all of the factors of $1+z$ together because
this form is more natural to implement in code.

@@

\noindent The limits of integration are technically from 0 to $\infty$, but these are
usually not written because the sensitivity $S(\lambda)$ is 0 for wavelengths
outside of a filter's bandpass.

@@

\noindent $P_x$ represents the zero-point for the filter $x$ on some particular
telescope. In order to use consistent magnitude values across telescopes that
have different light gathering abilities, we take the measured magnitude and
multiply it by the ratio of the standard flux rate to the flux rate for this
particular telescope and filter. For convenience, we use $P_x = -2.5
\text{log}(P'_x)$ so that we can work with flux instead of with magnitude.

Now that we have the identities in Equations \ref{eq:flux_definition} and
\ref{eq:flux2mag}, we will change directions and look at the definition of
K-correction $K_{xy}$. This value allows us to make an observation in filter
$y$ and report what the magnitude would have been in filter $x$ if no redshift
occurred:

@@

\noindent Note that this substitution omits the minus ($-$) sign that we used on a
similar substitution for $P_x$.

@@

Finally, we can use Equation \ref{eq:k_substitution} to convert $K'_{xy}$ back
into the magnitude $K_{xy}$:

@@

  This effect by itself can alter estimates of $H_0$ by several percent. While
  we present an estimate of the Hubble constant in Figure
  \ref{fig:H0bootstrap}, we are not terribly confident in the numbers because
  we were unable to correct for the zero-point errors, or verify that these
  errors may not exist.

@@

\section{Conclusions}
\label{sec:conslusions}

@@

As shown in Figure \ref{fig:mu_distance_vs_redshift}, when reported magnitudes
are corrected by adding the magnitude $-2.5\text{log}(1+z)$, there is a linear
relationship between redshift and luminosity distance. This is in accordance
with the Hubble-Lema\^{i}tre law. As opposed to the uncorrected values, there
is no visual acceleration. Since the corrected values are approximately linear,
we can use them to estimate the Hubble constant $H_0$. This is demonstrated in
Figure \ref{fig:expansion}. Furthermore, we performed a bootstrap calculation
to estimate the distribution of $H_0$ estimates in Figure
\ref{fig:H0bootstrap}.

\begin{figure}
  \plotone{bootstrapped_H0.pdf}
  \caption{A histogram of $10{,}000$ bootstrap trials measuring the Hubble constant
  $H_0$. Each trial samples an absolute magnitude $M \sim Norm(-19.2334,
  0.0404)$ for SnIa based on data published by \citet{camarena2020}. It then samples,
  with replacement, a population of supernovae from the dataset published by
  \citet{abbott2024}. Finally, it uses the non-parametric linear regression
  technique described by \citet{siegel1982}. The result of the bootstrap is $H_0
  \sim Norm(65.94, 1.3^2)$.
  }
\label{fig:H0bootstrap}
\end{figure}

The value estimated here, $H_0 = 65.94 \pm 1.3$ is highly suspicious
because of the flipped zero-point corrections, but it is consistent with
estimates of $H_0$ that are based on the CMB. \citet{planck2020} published the
CMB value of $H_0 = 67.27 \pm 0.6$, and the one sigma error bars overlap. It
is interesting to note that restricting the Dark Energy Survey observations of
SnIa \citep{vincenzi2024} to only those with $z < 0.1$, which presumably is a
small enough redshift that all observations will have been made with the same
filter, gives the estimate $H_0 = 67.8 \pm 1.9$.
